Question 4 

An American put option (right to sell a share) has a strike price of €100
with 3 periods left to maturity. The current share price is €105. Analysis
of the market indicates that the share price volatility per period is described
by: the stock will go up by €2 with probability 0.2, stay the same with
probability 0.3, decrease by €2 with probability 0.3 or decrease by €4 with
probability 0.2.
Use a Dynamic Programming formulation to investigate whether or not to
exercise the option, and if it is to be exercised when to exercise it. 25 %

Question 5 

(a) Define what is meant by a minimax strategy in a 2-player game. 3 %
(b) For the following matrix game
A B
A (5,3) (4,4)
B (2,6) (4,4)
Find the minimax strategies and payoffs for each player. If both players
play their minimax strategies, what is the value of the game ? 10 %
(c) In the context of a 2-person game, define what a Nash equilibrium is. 3 %
(d) Does the game of part (b) have Nash equilibria ? Justify your answer. 9 %


Question 6 

Consider the (symmetric) Cournot duopoly game: Firm i, i = 1, 2 produces
xi
items at a cost of
C(xi) = x
2
i
1000 + 3xi + 20
. The items sell at a price of
p(x1, x2) = 5 −
x1 + x2
500
each.
(a) Find the equilibrium of this game, and prove that it is a Nash equilibrium.
8 %
(b) Investigate this game if a collusive strategy is used. Contrast its solution
with that of part (a). 8 %
(c) If the game is to be played repeatedly, does it ever pay to defect from
the collusive strategy? In particular, consider the stern strategy: a firm
produces the collusive number of items until the other firm defects,
after which it reverts to producing the Cournot number of items. Using
the discount factor ω per period, when is this stern strategy a Nash
equilibrium ? 9 %


4 An American put option (right to sell a share) has a strike price of €100
with 3 periods left to maturity. The current share price is €102. Analysis
of the market indicates that the share price volatility per period is described
by: the stock will go up by €2 with probability 0.4, stay the same with
probability 0.3, decrease by €2 with probability 0.3.
Use a Dynamic Programming formulation to price the option. Should this
option be purchased for €0.3? Explain. 25
5 (a) In the context of a 2-person game, define what a Nash equilibrium is. 3
(b) What is a pure strategy ? What is a mixed strategy ? 3
(c) For strategies define strict dominance and weak dominance. 3
(d) By removing all strategies which are dominated by strict pure or mixed
strategies, derive the reduced version of the following 2-player matrix
game:
D E F
A (3,5) (5,1) (1,2)
B (1,1) (6,9) (6,4)
C (2,6) (4,7) (0,8)
8
(e) Derive the Nash equilibria and values of this game. 8
6 Consider the asymmetric duopoly game: Firm i, i = 1, 2 produces xi
items
at a cost of
C(xi) = 1
i
xi + 20.
The items sell at a price of
p(x1, x2) = 5 −
x1 + x2
500
each.
(a) Find the equilibrium of the game if it is played as a Cournot game,
and prove that it is a Nash equilibrium. 8
(b) Find the equilibrium if it played as a Stackelberg game with Firm 1 as
leader. 8
(c) Contrast and comment on the two solutions. 4
(d) Firm 1 receives an injection of capital and initiates a “leader strategy”.
However Firm 2 persists with its Cournot strategy. What happens and
what should each firm do in future interactions? 5


%=================================================================%

4 An American call option (right to buy a share) has a strike price of e100
with 3 periods left to maturity. The current share price is e98. Analysis
of the market indicates that the share price volatility per period is described
by: the stock will go up by e2 with probability 0.4, stay the same with
probability 0.3, decrease by e2 with probability 0.3.
Use a Dynamic Programming formulation to price the option. In particular,
if fi(x) is the expected value of the option at time i for a share price of x,
write down the Bellman equation for the scheme. 10
Find an expression for f2(x) as a function of x. 10
It can be shown that f0(98) = e0.544 and f0(102) = e2.87. Comment on
the significance of these. 5
5 (a) In the context of a 2-person game, define what a Nash equilibrium is. 3
(b) What is a pure strategy ? What is a mixed strategy ? 3
(c) For strategies define strict dominance and weak dominance. 3
(d) By removing all strategies which are dominated by strict pure or mixed
strategies, derive the reduced version of the following 2-player matrix
game:
D E F
A (4,-2) (3,0) (-3,-1)
B (-1,1) (2,2) (2,3)
C (2,1) (-1,-1) (0,4)
8
(e) Derive the Nash equilibria and values of this game. 8
6 (a) The costs incurred by a firm in a production period are
c = 100 + 2x
where x is the number of items produced in that period. The items sell
at a price of
p = 10 −
x
50
each. Find the level of production that maximises the firm’s profits
when the firm has a monopoly. 6
(b) If two identical firms supply the market with xi
, i = 1, 2 items each
at a cost per period of
ci = 100 + 2xi

respectively and sell each item at a price of
p = 10 −
x1 + x2
50 ,
analyse the resulting one shot Cournot game. 9
(c) If this 2-firm game is to be played repeatedly, consider the following
“cooperative” strategy: a firm produces half of the optimal level associated
with a monopoly (see part (a) for as long as the other firm does
the same, and if the other firm deviates, it reverts to the single shot
Cournot strategy thereafter. Does it ever pay to defect from the cooperative
strategy? In particular, using the discount factor ω per period,
when is this cooperative strategy a Nash equilibrium ? 10


1 (a) In terms of a strategic (matrix) game, what is a dominated strategy?
Describe the technique of iterated elimination of dominated strategies.
What is the technique used for ? 5
(b) By removing all strategies which are dominated by strict pure or mixed
strategies, derive the reduced version of the following 2-player zerosum
matrix game:
D E F
A (5,-5) (1,-1) (2,-2)
B (1,-1) (0,0) (3,-3)
C (2,-2) (3,-3) (6,-6)
8
(c) In terms of a 2-player zero-sum game, what is a minimax strategy? 3
(d) Derive the minimax strategies and value of the above game. 9
2 In a game show, contestants S´ile and Se´an start the last round with e500
and e400 respectively. Each must decide to pass or play. If a player passes,
they keep their money but if playing the first person gains e200 with probability
1/2 or loses all their money with probability 1/2, while the second
person gains or loses e200, each with probability 1/2. These outcomes are
independent of each other. The player with the most money at the end of
the round gets a bonus of e200.
(a) If S´ile goes first and Se´an sees her move, draw the game tree. 9
(b) Show that the strategic form of the game is
Pass Play
Pass (7,4) (6,5)
Play
9
2
, 5
9
2
, 5
where payoffs are expected values in 00’s. 8
(c) Solve the game. 8
3 Consider the following asymmetric duopoly game with isoelastic demand:
Two firms sell equivalent items at a price of
p(x1, x2) = 600
x1 + x2
per item, where Firm i,(i = 1, 2), produces xi
items at a cost of
C(xi) = mixi
.
The marginal costs are given by mi = i + 1 respectively.

(a) If the game is played as a Cournot game, show that the best response
xi = B(xj) of Firm i to Firm j is given by
B(xj) = r
600xj
mi
− xj
and hence find the Nash equilibrium. 8
(b) Find the equilibrium if it played as a Stackelberg game with Firm 1 as
leader. 8
(c) Contrast and comment on the two solutions. 4
(d) Firm 1 decides it will stick with the “leader strategy” in further production
cycles. However Firm 2 decides to ignore this and persist with
its Cournot strategy of part(a). What happens and what should each
firm do in future interactions?
