Question 4 

An American put option (right to sell a share) has a strike price of €100
with 3 periods left to maturity. The current share price is €105. Analysis
of the market indicates that the share price volatility per period is described
by: the stock will go up by €2 with probability 0.2, stay the same with
probability 0.3, decrease by €2 with probability 0.3 or decrease by €4 with
probability 0.2.
Use a Dynamic Programming formulation to investigate whether or not to
exercise the option, and if it is to be exercised when to exercise it. 25 %

Question 5 

(a) Define what is meant by a minimax strategy in a 2-player game. 3 %
(b) For the following matrix game
A B
A (5,3) (4,4)
B (2,6) (4,4)
Find the minimax strategies and payoffs for each player. If both players
play their minimax strategies, what is the value of the game ? 10 %
(c) In the context of a 2-person game, define what a Nash equilibrium is. 3 %
(d) Does the game of part (b) have Nash equilibria ? Justify your answer. 9 %


Question 6 

Consider the (symmetric) Cournot duopoly game: Firm i, i = 1, 2 produces
xi
items at a cost of
C(xi) = x
2
i
1000 + 3xi + 20
. The items sell at a price of
p(x1, x2) = 5 −
x1 + x2
500
each.
(a) Find the equilibrium of this game, and prove that it is a Nash equilibrium.
8 %
(b) Investigate this game if a collusive strategy is used. Contrast its solution
with that of part (a). 8 %
(c) If the game is to be played repeatedly, does it ever pay to defect from
the collusive strategy? In particular, consider the stern strategy: a firm
produces the collusive number of items until the other firm defects,
after which it reverts to producing the Cournot number of items. Using
the discount factor ω per period, when is this stern strategy a Nash
equilibrium ? 9 %
