Thomas S. Ferguson
Part Player 2. Two-Person Zero-Sum Games
1. The Strategic Form of a Game.
1.1 Strategic Form.
1.2 Example: Odd or Even.
1.3 Pure Strategies and Mixed Strategies.
1.4 The Minimax Theorem.
1.5 Exercises.
2. Matrix Games. Domination.
2.1 Saddle Points.
2.2 Solution of All 2 by 2 Matrix Games.
2.3 Removing Dominated Strategies.
2.4 Solving 2 × n and m × 2 Games.
2.5 Latin Square Games.
2.6 Exercises.
3. The Principle of Indifference.
3.1 The Equilibrium Theorem.
3.2 Nonsingular Game Matrices.
3.3 Diagonal Games.
3.4 Triangular Games.
3.5 Symmetric Games.
3.6 Invariance.
3.7 Exercises.
Player 2 – 1
4. Solving Finite Games.
4.1 Best Responses.
4.2 Upper and Lower Values of a Game.
4.3 Invariance Under Change of Location and Scale.
4.4 Reduction to a Linear Programming Problem.
4.5 Description of the Pivot Method for Solving Games.
4.6 A Numerical Example.
4.7 Approximating the Solution: Fictitious Play.
4.8 Exercises.
5. The Extensive Form of a Game.
5.1 The Game Tree.
5.2 Basic Endgame in Poker.
5.3 The Kuhn Tree.
5.4 The Representation of a Strategic Form Game in Extensive Form.
5.5 Reduction of a Game in Extensive Form to Strategic Form.
5.6 Example.
5.7 Games of Perfect Information.
5.8 Behavioral Strategies.
5.9 Exercises.
6. Recursive and Stochastic Games.
6.1 Matrix Games with Games as Components.
6.2 Multistage Games.
6.3 Recursive Games. -Optimal Strategies.
6.4 Stochastic Movement Among Games.
6.5 Stochastic Games.
6.6 Approximating the Solution.
6.7 Exercises.
7. Infinite Games.
7.1 The Minimax Theorem for Semi-Finite Games.
Player 2 – 2
7.2 Continuous Games.
7.3 Concave and Convex Games.
7.4 Solving Games.
7.5 Uniform[0,1] Poker Models.
7.6 Exercises.
References.
Player 2 – 3
Part Player 2. Two-Person Zero-Sum Games

