II – 93
References.
Robert J. Aumann and Michael B. Maschler (1995) Repeated Games of Incomplete Information,
The MIT Press, Cambridge, Mass.
A. Ba˜nos (1968) On pseudo-games. Ann. Math. Statist. 39, 1932-1945.
V. J. Baston, F. A. Bostock and T. S. Ferguson (1989) The number hides game, Proc.
Amer. Math. Soc. 107, 437-447.
John D. Beasley (1990) The Mathematics of Games, Oxford University Press.
Emile Borel (1938) Trait´ ´ e du Calcul des Probabilit´es et ses Applications Volume IV, Fascicule
2, Applications aux jeux des hazard, Gautier-Villars, Paris.
G. W. Brown (1951) ”Iterative Solutions of Games by Fictitious Play” in Activity Analysis
of Production and Allocation, T.C. Koopmans (Ed.), New York: Wiley.
G. S. Call and D. J. Velleman (1993) Pascal’s matrices, Amer. Math. Mo. 100, 372-376.
M. T. Carroll, M. A. Jones, E. K.Rykken (2001) The Wallet Paradox Revisited, Math.
Mag. 74, 378-383.
W. H. Cutler (1975) An optimal strategy for pot-limit poker, Amer. Math. Monthly 82,
368-376.
W. H. Cutler (1976) End-Game Poker, Preprint.
Melvin Dresher (1961) Games of Strategy: Theory and Applications, Prentice Hall, Inc.
N.J.
Melvin Dresher (1962) A sampling inspection problem in arms control agreements: a gametheoretic
analysis, Memorandum RM-2972-ARPA, The RAND Corporation, Santa
Monica, California.
R. J. Evans (1979) Silverman’s game on intervals, Amer. Math. Mo. 86, 277-281.
H. Everett (1957) Recursive games, Contrib. Theor. Games III, Ann. Math. Studies 39,
Princeton Univ. Press, 47-78.
C. Ferguson and T. Ferguson (2007) The Endgame in Poker, in Optimal Play: Mathematical
Studies of Games and Gambling, Stuart Ethier and William Eadington, eds.,
Institute for the Study of Gambling and Commercial Gaming, 79-106.
T. S. Ferguson (1967) Mathematical Statistics — A Decision-Theoretic Approach, Academic
Press, New York.
T. S. Ferguson and C. Melolidakis (1997) Last Round Betting, J. Applied Probability 34
974-987.
J. Filar and K. Vrieze (1997) Competitive Markov Decision Processes, Springer-Verlag,
New York.
L. Friedman (1971) Optimal bluffing strategies in poker, Man. Sci. 17, B764-B771.
II – 94
S. Gal (1974) A discrete search game, SIAM J. Appl. Math. 27, 641-648.
M. Gardner (1978) Mathematical Magic Show, Vintage Books, Random House, New York.
Andrey Garnaev (2000) Search Games and Other Applications of Game Theory, Lecture
Notes in Economics and Mathematical Systems 485, Springer.
G. A. Heuer and U. Leopold-Wildburger (1991) Balanced Silverman Games on General
Discrete Sets, Lecture Notes in Econ. & Math. Syst., No. 365, Springer-Verlag.
R. Isaacs (1955) A card game with bluffing, Amer. Math. Mo. 62, 99-108.
S. M. Johnson (1964) A search game, Advances in Game Theory, Ann. Math. Studies 52,
Princeton Univ. Press, 39-48.
Samuel Karlin (1959) Mathematical Methods and Theory in Games, Programming and
Economics, in two vols., Reprinted 1992, Dover Publications Inc., New York.
H. W. Kuhn, (1950) A simplified two-person poker, Contrib. Theor. Games I, Ann. Math.
Studies 24, Princeton Univ. Press, 97-103.
H. W. Kuhn (1997) Classics in Game Theory, Princeton University Press.
A. Maitra and W. Sudderth (1996) Discrete Gambling and Stochastic Games, in the Series
Applications in Mathematics 32, Springer.
J. J. C. McKinsey (1952) Introduction to the Theory of Games, McGraw-Hill, New York.
N. S. Mendelsohn (1946) A psychological game, Amer. Math. Mo. 53, 86-88.
J. Milnor and L. S. Shapley (1957) On games of survival, Contrib. Theor. Games III, Ann.
Math. Studies 39, Princeton Univ. Press, 15-45.
Kent E. Morrison (2010) The Multiplication Game, Math. Mag. 83, 100-110.
J. F. Nash and L. S. Shapley (1950) A simple 3-person poker game, Contrib. Theor. Games
I, Ann. Math. Studies 24, Princeton Univ. Press, 105-116.
D. J. Newman (1959) A model for “real” poker, Oper. Res. 7, 557-560.
Guillermo Owen (1982) Game Theory, 2nd Edition, Academic Press.
T. E. S. Raghavan, T. S. Ferguson, T. Parthasarathy and O. J. Vrieze, eds. (1991) Stochastic
Games and Related Topics, Kluwer Academic Publishers.
J. Robinson (1951) An Iterative Method of Solving a Game, Annals of Mathematics 54,
296-301.
W. H. Ruckle (1983) Geometric games and their applications, Research Notes in Mathematics
82, Pitman Publishing Inc.
L. S. Shapley (1953) Stochastic Games, Proc. Nat. Acad. Sci. 39, 1095-1100.
L. S. Shapley and R. N. Snow (1950) Basic solutions of discrete games, Contrib. Theor.
Games I, Ann. Math. Studies 24, Princeton Univ. Press, 27-35.
II – 95
S. Sorin and J. P. Ponssard (1980) The LP formulation of finite zero-sum games with
incomplete information, Int. J. Game Theory 9, 99-105.
Philip D. Straffin (1993) Game Theory and Strategy, Mathematical Association of America.
John Tukey (1949) A problem in strategy, Econometrica 17, 73.
J. von Neumann and O. Morgenstern (1944) The Theory of Games and Economic Behavior,
Princeton University Press.
J. D. Williams, (1966) The Compleat Strategyst, 2nd Edition, McGraw-Hill, New York.
II – 96
